Are we learning to build better models at the cost of better decisions?

### 2021-03-10 manuscript/manuscript.Rmd
:pen:

# Discussion
<!--
hit bigger themes first & objections first
- Learning can be counter-productive
- Value of information can be negative in practice

Objections:
- we already know simple models can be better, overfit models can lead us astray
- we already know Knightian uncertainty matters.  (underestimating uncertainty isn't the same thing)

- is this the result of underestimating the uncertainty?

Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Appendix B, Fig S1).

@@

Note that the mechanism shown here has nothing to do with the much more familiar issue of over-fitting, which also involves a better-fitting model leading to worse outcomes [].  In fact, an model which has been over-fit will gradually be rejected by either the iterative forecasting or adaptive management approaches shown here, as these approaches continually confront the models with new data to which they had not been previously fit.  The ability to avoid over-fitting is one of the greatest appeals of any iterative management strategy. But in the example here, both Model 1 and Model 2 have the same structural complexity.  The better decision performance of Model 1 is a consequence of capturing the key attribute needed for a good decision, which is perhaps a surprisingly an all-together different criteria than fit.

Previous work has acknowledged that iterative learning may not always be feasible and that it may not always be beneficial, but has failed to recognize the possibility that it can potentially be detrimental.  For example, active adaptive management is premised on the observation that future observations may be too uninformative to distinguish between alternatives [@Walters1978].  Others have also noted that even big improvements in predictive accuracy could have negligible improvement on the decision outcomes, which motivates quantifying the Value of Information (VOI) [@Katz1987]. But in both cases, the worse outcome is wasted effort.  In the scenario considered here, the VOI is negative: any learning over the proposed model uncertainty leads to lower expected net utility than not learning.  The only way to avoid this trap is to _forego learning_ with each new observation, to avoid iterative updates. A manager must follow the decisions proposed by model 1 for many iterations to realize that it leads to more desirable outcomes than alternative models.  This raises several questions.  Do such methods of learning and reducing model uncertainty ever lead us towards worse long-term outcomes in the real world? If so, How would we know, and what should we do about it? I address these in turn.

@@

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  I contrast that position with two common but diametrically opposed objections: (1) that the issue is easily resolved by including more or better models, or (2) this issue is easily resolved by relying less on models.  These are both important points.

The union of two simple models considered here (Models 1 & 2) certainly fails to fully capture the inherent uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood [].  Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations.   However, it must also be noted that we could have included a much more realistic approach of uncertainty over many possible values of $r$, $K$ (see Appendix B, Fig S1), or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], without altering the results at all (these models also all have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  Surely it would be preferable if our methods could prefer the nearly optimally-performing Model 1 which we already have than require the issue be solved only by adding more and more models?

If we cannot always resolve the issue by adding more models, is this just another example of the risks of an over-reliance on models at all?

<!-- (3) THIS problem (4) THIS problem. I like that you outline the points you are going to cover! -->

>>>>

# Discussion

<!-- 
First: Intuition for the result.
  - See that model 1 captures the key feature, while model 2 is just 'closer to the curve'
  - See that this is a generic truth about decision-making: there's always a Model 1
  - This is not about over-fitting.

-->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Appendix B, Fig S1).

@@

Note that the mechanism shown here has nothing to do with the much more familiar issue of over-fitting, in which a better-fitting model will also lead to worse outcomes [].  In fact, an model which has been over-fit will gradually be rejected by either the iterative forecasting or adaptive management approaches shown here, as these approaches continually confront the models with new data to which they had not been previously fit.  The ability to avoid over-fitting is one of the greatest appeals of any iterative management strategy. But in the example here, both Model 1 and Model 2 have the same structural complexity.  The better decision performance of Model 1 is a consequence of capturing the key attribute needed for a good decision, which is perhaps a surprisingly an all-together different criteria than fit.

This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood []. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ (see Appendix B, Fig S1), or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

<!--
So what?

- This is different than previous work: e.g. VOI is not zero, it's negative here.
- Model 1 is not a unicorn: we cannot rule out these situations in real world (particularly ecology where we don't actually do much iterative updating in practice)
- 
-->

Previous work has acknowledged that iterative learning may not always be feasible and that it may not always be beneficial, but has failed to recognize the possibility that it can potentially be detrimental.  For example, active adaptive management is premised on the observation that future observations may be too uninformative to distinguish between alternatives [@Walters1978].  Others have also noted that even big improvements in predictive accuracy could have negligible improvement on the decision outcomes, which motivates quantifying the Value of Information (VOI) [@Katz1987]. But in both cases, the worse outcome is wasted effort.  In the scenario considered here, the VOI is negative: any learning over the proposed model uncertainty leads to lower expected net utility than not learning.

The only way to avoid this trap is to _forego learning_ with each new observation, to avoid iterative updates.  It may seem that the adaptive management approach fails because posterior probabilities are updated according to Bayes rule. Like iterative forecasting, this favors models which better predict the data, despite the fact that under the adaptive management approach, the overall optimization is conditioned on actual utility and not model fit.  However, this issue is not easily avoided.  For example, so-called greedy optimization techniques that favor actions with higher immediate reward do even worse in this context, since such an algorithm would obviously maximize the immediate harvest and therefore collapse the stock.  Only by comparing the net utility derived from management under model 1 for many iterations to the net utility derived under model 2 after many iterations can we deterime that model 1 leads to better outcomes.  This raises several questions.  Do such methods of learning and reducing model uncertainty ever lead us towards worse long-term outcomes in the real world? If so, How would we know, and what should we do about it?

@@

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  (1) that the issue is just an example of the already-understood problem of underspecifying model uncertainty, which is resolved by including more or better models, or (2) this issue is just an example of the already-understood limitations of relying too much on models to shape decision-making. Either we need more models, or we need fewer models.

If to advocates of iterative model updating approaches the solution is always more models, then to skeptics the solution is fewer models and algorithms.  
If we cannot always resolve the issue by adding more models, is this just another example of the risks of an over-reliance on models at all?
